Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-28 fmics/paper.tex
incorporate Naghmeh's updates

To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
%Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
%Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
Our central result equates the output of a slice to a sum of consecutive input counts.
More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
\begin{theorem}

The output of slice n at update time t ... (statement of the theorem)
\begin{multline}\label{output_input_at_update_times}
\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
\end{multline}
\end{theorem}

\begin{proof}
\begin{enumerate}
\item
$\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)$, since shift register cells get their values from the previous cell, but only update every $\mathsf{delay}\;n$ time steps;
\item
$\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))$, by induction on $m$ using the result above;
\item
$\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), which can be proved by induction on $t$;
\item
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$;
\item
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output;
\item
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
\end{enumerate}

\end{proof}

>>>>

%Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
Figure~\ref{fig:auxdefs} defines a set of additional functions required to prove our main results.
The function $\mathsf{delay\_sum}\;n$ represents the accumulative delay of the preceding shift registers of a slice.
For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated and the $\mathsf{exact}\;D\;n\;x\;t$ {\bf Ramana, can say in simple words what exact represents?}

%Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
Our central result equates the output of a slice to a sum of consecutive input counts.
More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
Note that the precise statement of such a theorem is complicated by details of the value of the $\mathsf{output}$ at times when the shift registers have not yet received any values.

%To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}

\begin{theorem}
For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at update time $t$ is equal to
\begin{equation*}%\label{output_input_at_update_times}
%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0 \qquad\qquad\qquad  t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
\end{equation*}
\end{theorem}

\begin{proof}
The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore,
\begin{equation*}
\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)
\end{equation*}
By induction on $m$ on the above,
\begin{equation*}
\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))
\end{equation*}
By induction on $t$, we can show that the $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
\begin{equation*}
\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
\end{equation*}
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$.
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output.
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
\qed
\end{proof}
